### 2018-11-19 docs/slides/img/script.Rmd
reinstate docs/slides

# code to make figures

```{r}
source ("../script.R")
promenade <- readRDS ("../data/osm-promenade.Rds")
ms48155 <- readRDS ("../data/osm-ms-48155.Rds")
ms <- readRDS ("../data/osm-ms-highways.Rds")
```

## mapview images

```{r}
mapview::mapview (ms48155$osm_lines)
```

```{r}
mapview::mapview (promenade$osm_lines)
```

## Mapdeck images

These are made with mapdeck which doesn't currently integrate with xaringan, so
have to be manually screenshotted.

`img/ms48155lines-mapview.png`:
```{r mapdeck1, eval = FALSE, echo = FALSE}
library (osmdata)
library (dodgr)
library (mapdeck)
loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
tok <- Sys.getenv ("MAPBOX_TOKEN")
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80)
```

`img/mspromenade-mapview.png`:
```{r mapdeck2, eval = FALSE, echo = FALSE}
library (dodgr)
loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
tok <- Sys.getenv ("MAPBOX_TOKEN")
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = promenade$osm_lines,
              layer_id = "promenade",
              stroke_width = 5,
              stroke_colour = "#33cc33",
              stroke_opacity = 100)
```

`img/ms-highway-promenade-mapview.png`:
```{r mapdeck2, eval = FALSE, echo = FALSE}
library (dodgr)
loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
tok <- Sys.getenv ("MAPBOX_TOKEN")
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80) %>%
    add_path (data = promenade$osm_lines,
              layer_id = "promenade",
              stroke_width = 5,
              stroke_colour = "#33cc33",
              stroke_opacity = 100)
```

# ms components

```{r}
library (dodgr)
g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         pitch = 0,
         zoom = 12,
         location = loc) %>%
    add_line (data = g,
              layer_id = "ms-highways",
              origin = c("from_lon", "from_lat"),
              destination = c("to_lon", "to_lat"),
              stroke_colour = "component",
              stroke_width = 1,
              palette = rainbow)
```

`img/ms-random-flows.png`:
```{r}
f <- readRDS ("data/dodgr-flows-ms.Rds")
f <- dodgr::merge_directed_flows (f)
library (mapdeck)
f$flow <- 20 * f$flow / max (f$flow)
pal <- colorRampPalette (c ("lawngreen", "red"))
loc <-  c (7.619786, 51.952610)
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         pitch = 0,
         zoom = 12,
         location = loc) %>%
    add_line (data = f,
              layer_id = "ms-highways",
              origin = c("from_lon", "from_lat"),
              destination = c("to_lon", "to_lat"),
              stroke_colour = "flow",
              stroke_width = "flow",
              palette = pal)
```

`img/ms-random-flows90.png`:
```{r}
ms <- readRDS ("data/osm-ms-highways.Rds")
library (osmdata)
library (dodgr)
ms <- osm_poly2line (ms)
# get OSM way IDs for promenade
index <- grep ("promenade", ms$osm_lines$name, ignore.case = TRUE)
kp <- grep ("kanal", ms$osm_lines$name, ignore.case = TRUE)
index <- index [!index %in% kp]
ids <- ms$osm_lines$osm_id [index] %>% as.character ()
index <- which (g$way_id %in% ids)

v <- dodgr_vertices (g)
pts <- sample (v$id, size = 1000)
flowmat <- matrix (1, nrow = 1000, ncol = 1000)

g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
g$d_weighted [index] <- g$d [index] * 0.9
f <- dodgr_flows_aggregate (g, from = pts, to = pts, flows = flowmat)
saveRDS (f, file = "data/dodgr-flows-ms90.Rds")

g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
g$d_weighted [index] <- g$d [index] * 0.8
f <- dodgr_flows_aggregate (g, from = pts, to = pts, flows = flowmat)
saveRDS (f, file = "data/dodgr-flows-ms80.Rds")

g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
g$d_weighted [index] <- g$d [index] * 0.7
f <- dodgr_flows_aggregate (g, from = pts, to = pts, flows = flowmat)
saveRDS (f, file = "data/dodgr-flows-ms70.Rds")
```
```{r}
plot1map <- function (f)
{
    if (is.character (f)) # name of file
    {
        f <- readRDS (f)
        f <- dodgr::merge_directed_flows (f)
    }
    library (mapdeck)
    f$flow <- 20 * f$flow / max (f$flow)
    pal <- colorRampPalette (c ("lawngreen", "red"))
    loc <-  c (7.619786, 51.962)
    mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
             pitch = 0,
             zoom = 14,
             location = loc) %>%
    add_line (data = f,
              layer_id = "ms-highways",
              origin = c("from_lon", "from_lat"),
              destination = c("to_lon", "to_lat"),
              stroke_colour = "flow",
              stroke_width = "flow",
              palette = pal)
}
plot1map ("data/dodgr-flows-ms.Rds")
plot1map ("data/dodgr-flows-ms90.Rds")
plot1map ("data/dodgr-flows-ms80.Rds")
plot1map ("data/dodgr-flows-ms70.Rds")
```

# exposure

## vehicular flows

Same adjustment has to be made to the Ring, which is "secondary", and to lots of
the inner-city roads which are "primary" to get flows concentrated around the
ring rather than through the middle
```{r}
library (osmdata)
ms <- readRDS ("data/osm-ms-highways.Rds")
ms <- osm_poly2line (ms)

# get OSM way IDs for Ring, which are all secondary
index_ring <- which (grepl ("ring", ms$osm_lines$name, ignore.case = TRUE) &
                ms$osm_lines$highway == "secondary")
index_sg <- which (grepl ("am stadtgraben", ms$osm_lines$name, ignore.case = TRUE) &
                   ms$osm_lines$highway == "primary")
index_sp <- which (grepl ("schlossplatz", ms$osm_lines$name, ignore.case = TRUE) &
                   ms$osm_lines$highway == "primary")
index_nt <- which (grepl ("neutor", ms$osm_lines$name, ignore.case = TRUE) &
                   ms$osm_lines$highway == "primary")
index_st <- which (grepl ("steinfurter", ms$osm_lines$name, ignore.case = TRUE) &
                   ms$osm_lines$highway == "primary")
index_s <- unique (c (index_sg, index_sp, index_nt, index_st))
ids_ring <- ms$osm_lines$osm_id [index_ring]
ids_s <- ms$osm_lines$osm_id [unique (c (index_sg, index_sp, index_nt, index_st))]
```

```{r}
weighting_profiles [weighting_profiles$name == "motorcar", ]
```
wt_profile for secondary is 0.8, and primary is 0.9

functions from `flowlayers`:
```{r}
dist_to_lonlat_range <- function (verts, d = 20)
{
    xy0 <- c (mean (verts$x), mean (verts$y))
    names (xy0) <- c ("x", "y")
    minf <- function (a, xy0) { abs (geodist::geodist (xy0, xy0 + a) - d) }
    optimise (minf, c (0, 0.1), xy0)$minimum
}

disperse_flows <- function (graph, d = 20)
{
    # This actually works by mapping flows to vertices, and then applying a
    # Gaussian kernel to those points, before re-mapping them back to edges
    v <- dodgr::dodgr_vertices (graph)
    sig <- dist_to_lonlat_range (v, d)

    indxf <- match (v$id, graph$from_id)
    indxt <- match (v$id, graph$to_id)
    indxf [is.na (indxf)] <- indxt [is.na (indxf)]
    indxt [is.na (indxt)] <- indxf [is.na (indxt)]
    v$flow <- apply (cbind (graph$flow [indxf], graph$flow [indxt]), 1, max)

    # warnings are issued here if any (x, y) are duplicated
    xy <- suppressWarnings (spatstat::ppp (v$x, v$y, range (v$x), range (v$y)))
    d <- spatstat::density.ppp (xy, weights = v$flow, sigma = sig, at = "points")
    vsum <- sum (v$flow)
    d <- d * vsum / sum (d)
    indx <- which (d > v$flow)
    v$flow [indx] <- d [indx]
    v$flow <- v$flow * vsum / sum (v$flow)

    # Then map those vertex values back onto the graph
    indxf <- match (graph$from_id, v$id)
    indxt <- match (graph$to_id, v$id)
    fmax <- apply (cbind (v$flow [indxf], v$flow [indxt]), 1, max)
    fsum <- sum (graph$flow)
    fmax <- fmax * fsum / sum (graph$flow)
    indx <- which (fmax > graph$flow)
    # That again imbalances the flow of the graph, so needs to be standardised
    # back to original sum again
    fsum <- sum (graph$flow)
    graph$flow [indx] <- fmax [indx]
    graph$flow <- graph$flow * fsum / sum (graph$flow)

    return (graph)
}
```

```{r}
library (dodgr)
gv <- weight_streetnet (ms$osm_lines, wt_profile = "motorcar")
rrr <- 4
index <- match (ids_ring, gv$way_id)
gv$d_weighted [index] <- gv$d [index] / rrr
index <- match (ids_s, gv$way_id)
gv$d_weighted [index] <- gv$d [index] * rrr
gc <- dodgr_contract_graph (gv)
v <- dodgr_vertices (gc$graph)
pts <- sample (v$id, size = 1000)
message ("aggregating ... ", appendLF = FALSE)
fv <- dodgr_flows_aggregate (gv, from = pts, to = pts,
                             flow = matrix (1, nrow = 1000, ncol = 1000))
message ("done\ndispersing ... ", appendLF = FALSE)
fvd <- disperse_flows (fv, 20) %>%
    merge_directed_flows ()
fv <- merge_directed_flows (fv)
mesage ("done")
```
The two maps of vehicular flows:
```{r}
plot1map (fv)
plot1map (fvd)
```
```{r}
f <- readRDS ("data/dodgr-flows-ms70.Rds") %>%
    merge_directed_flows ()
f <- f [which (f$edge_id %in% fvd$edge_id), ]
indx1 <- match (f$edge_id, fvd$edge_id)
indx2 <- match (fvd$edge_id [indx1], f$edge_id)
f$flow <- f$flow / max (f$flow)
fvd$flow <- fvd$flow / max (fvd$flow)
f$flow <- f$flow * fvd$flow [indx1] [indx2]
plot1map (f)
```