### 2018-11-20 ms-meetup-nov18.Rmd
:rocket:

---
title: "Using R to help plan the <br> future of transport"
subtitle: "Münster R Users Group"
author: "Mark Padgham <br> Active Transport Futures"
date: "Tuesday 20th November, 2018 "
output:
      xaringan::moon_reader:
            yolo:
                        img: img/fortune-beach.png
                        times: 2
            lib_dir: libs
            css: xaringan-themer.css
            nature:
                  highlightStyle: github
                  highlightLines: true
                  countIncrementalSlides: false
---
```{r load, include = FALSE, eval = TRUE}
library(xaringanthemer)
source ("myxaringantheme.R")
source ("data-script.R")
promenade <- readRDS ("data/osm-promenade.Rds")
ms48155 <- readRDS ("data/osm-ms-48155.Rds")
net <- f <- readRDS ("data/dodgr-flows-ms.Rds")
net$flow <- NULL
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)# suppress version num in subdir name

knitr::opts_chunk$set(cache = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

class: left, middle, inverse

.left-column[
`r icon::fa_github(size = 2, animate = "spin")` atfutures

`r icon::fa_github(size = 2, animate = "spin")` mpadge

`r icon::fa_github(size = 2, animate = "spin")` ropensci
]

.right-column[
`r icon::fa_twitter(size = 2, animate = "spin")` bikesRdata

`r icon::fa_envelope(size = 1)` .small[mark.padgham@email.com]<br><br>

`r icon::ii_ios_world_outline(size = 5)` atfutures.github.io
]

.box-bottom[
slides at <br>
[https://github.com/mpadge/ms-user-meetup-nov2018](https://github.com/mpadge/ms-user-meetup-nov2018)
]

---
class: left, top, inverse

### Primary Collaborators

- Robin Lovelace (@robinlovelace)
- Layik Hama (@layik)
- Andreas Petutschnig (@karpfen)
- Andrew Smith (@virgesmith)

### Additional `r icon::fa_github(animate = "spin")` Contributions From ...

.left-column[
- RichardEllison
- tbuckl
- skcott
- hrbrmstr
- maelle
- mdsumner
- mem48
- jimshady
- angela-li
  ]

.right-column[
- graceli8
- arfon
- jimhester
- tazinho
- karthik
- fzenoni
- mhenderson
- patperu
- eyesofbambi
  ]

---
class: left, top, inverse

# Active Transport Futures

---
background-image: url(img/amsterdam-cars.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: left, top, inverse

# Active Transport Futures

---
class: left, top, inverse
background-image: url(img/ms-promenade.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

# Active Transport Futures

---
class: left, top, inverse
background-image: url(img/kl-bike-path.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

# [Active Transport Futures](http://www.who.int/sustainable-development/cities/en/)

---

# Active Transport Futures

For a given city:

- Collate data on patterns of urban mobility
- Collate data on transport networks <br> (road, pedestrian, public transport)
- Analyse "flows" throughout the network(s)
- Collate data on (spatial patterns of) air pollution
- Analyse dynamic interactions between movement, air pollution, and health
- Develop, compare, contrast scenarios for ameliorating negative effects while
  enhancing positive effects
- [Health Economic Assessment Tool (HEAT)](https://heatwalkingcycling.org/#homepage)

---

# Active Transport Futures

For a given city:

- ~~Collate data on patterns of urban mobility~~
- Collate data on transport networks <br> (road, pedestrian, public transport)
- Analyse "flows" throughout the network(s)
- ~~Collate data on (spatial patterns of) air pollution~~
- Analyse dynamic interactions between movement, air pollution~~, and health~~
- ~~Develop, compare, contrast scenarios for ameliorating negative effects while
  enhancing positive effects~~
- ~~[Health Economic Assessment Tool (HEAT)](https://heatwalkingcycling.org/#homepage)~~

---

# Active Transport Futures

## Software: R Packages

- `osmdata` (Open Street Map data)
- `dodgr` (Distances On Directed GRaphs)
- `stplanr` (Sustainable Transport PLANning in R)
- `osmplotr` (OSM plotting routines)
- `geoplumber` (Turbo-charged web server for geo data)
- `bikedata` (Data from public hire bicycles)
- `spatialcluster`
- ... plus more

- demo of `geoplumber` at [`http://35.233.61.182`](http://35.233.61.182)

---

# Active Transport Futures

## Software: R Packages

- `osmdata` (Open Street Map data)
- `dodgr` (Distances On Directed GRaphs)
- ~~`stplanr` (Sustainable Transport PLANning in R)~~
- ~~`osmplotr` (OSM plotting routines)~~
- ~~`geoplumber` (Turbo-charged web server for geo data)~~
- ~~`bikedata` (Data from public hire bicycles)~~
- ~~`spatialcluster`~~
- ... plus more

- demo of `geoplumber` at [`http://35.233.61.182`](http://35.233.61.182)

---
class: left, top

# The osmdata package

OSM = Open Street Map

- CRAN
- `r icon::fa_github(size = 1, animate = "spin")` `ropensci/osmdata`
- Several mentions on
  [OSM wiki](https://wiki.openstreetmap.org/wiki/OSM_Scientific_Tools)

--

## Why OSM data?

---
class: center, top, inverse
.fontBIG[
Popquiz: <br><br>What is the least mapped country on the planet?
]

---
class: center, middle, inverse
background-image: url(img/osm-north-korea.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---
class: center, middle, inverse
background-image: url(img/anju-bing.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---
class: center, middle, inverse
background-image: url(img/anju-google.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---
class: center, middle, inverse
background-image: url(img/anju-google2.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---
class: center, middle, inverse
background-image: url(img/anju-osm.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---

# The osmdata package

- CRAN
- `r icon::fa_github(size = 1, animate = "spin")` `ropensci/osmdata`
- Several mentions on
  [OSM wiki](https://wiki.openstreetmap.org/wiki/OSM_Scientific_Tools)

## Why OSM data?

--

Because Open Street Map is the most comprehensive street atlas humanity has ever
created.

## How?

`osmdata` uses the `overpass` server to retrieve data

---

# The osmdata package

```{r opq1}
library (osmdata)
opq ("muenster de")
```

---

# The osmdata package

```{r opq2}
library (osmdata)
opq ("48155 de")
```

---

# The osmdata package

```{r add-osm-ft}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway")
```

---

# The osmdata package

```{r osmdata-sf, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
```
```{r osmdata-sf-fakey, eval = TRUE, echo = FALSE, message = TRUE}
print (ms48155)
```

---

# The osmdata package

```{r osmdata-sf2, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
```

- `opq` `r icon::fa_long_arrow_alt_right()` .bold[WHERE]
- `add_osm_feature` `r icon::fa_long_arrow_alt_right()` .bold[WHAT]
- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[HOW]

---

# The osmdata package - Spatial Formats

```{r osmdata-sf3, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
```

- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[Simple Features]

---

# The osmdata package - Spatial Formats

```{r osmdata-pbf, eval = FALSE}
library (osmdata)
opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf (file = "ms48155.pbf")
```

- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[Simple Features]
- `osmdata_sp` `r icon::fa_long_arrow_alt_right()` .bold[R Spatial]
- `osmdata_sc` `r icon::fa_long_arrow_alt_right()` .bold[Silicate (Simplicial Complex)]
- `osmdata_xml` `r icon::fa_long_arrow_alt_right()` .bold[XML / OSM]
- `osmdata_pbf` `r icon::fa_long_arrow_alt_right()` .bold[PBF (Protocol Buffer Format)]

---

# The osmdata package

```{r osmdata_sf2, eval = FALSE, message = TRUE}
library (osmdata)
ms48155 <- opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
ms48155
```
```{r osmdata_sf2-fakey, echo = FALSE, message = TRUE}
print (ms48155)
```
---

# The osmdata package - Visualisation

```{r mapview, eval = FALSE}
library (osmdata)
ms48155 <- opq ("48155 de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
library (mapview)
mapview (ms48155$osm_lines)
```

---
background-image: url(img/ms48155lines-mapview.png)
background-size: contain 
background-position: 50% 50%
class: center, top, inverse

---

# The osmdata package - Visualisation

How to visualise just the Promenade?

```{r mapview2, eval = FALSE}
library (osmdata)
ms48155 <- opq ("48155 de") %>% #<<
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
```
---

# The osmdata package - Visualisation

How to visualise just the Promenade?

```{r promenade-data1, eval = FALSE, message = TRUE}
library (osmdata)
promenade <- opq ("münster de") %>% #<<
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
```

---

# The osmdata package - Visualisation

How to visualise just the Promenade?

```{r promenade-data2, eval = FALSE}
library (osmdata)
promenade <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    add_osm_feature (key = "name", value = "promenade", #<<
                     value_exact = FALSE) %>%           #<<
    osmdata_sf ()
```

---

# The osmdata package - Visualisation

How to visualise just the Promenade?

```{r promenade-data3, eval = FALSE}
library (osmdata)
promenade <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    add_osm_feature (key = "name", value = "promenade",
                     value_exact = FALSE) %>%
    osmdata_sf ()
library (mapview)
mapview (promenade$osm_lines) #<<
```

---
background-image: url(img/mspromenade-mapview.png)
background-size: contain 
background-position: 50% 50%
class: center, middle, inverse

---

class: middle, inverse, center

.fonthuge[
Digression:<br>Visualising OSM data
]

---

# Visualising osmdata

```{r, eval = FALSE}
library (mapview)
mapview (ms48155$osm_lines)
```

---
background-image: url(img/ms48155lines-mapview.png)
background-size: contain 
background-position: 50% 50%
class: center, middle, inverse

---

# Visualising osmdata

```{r, eval = FALSE}
library (mapview)
mapview (ms48155$osm_lines) %>%
    addFeatures (promenade$osm_lines)
```
... currently will not work (coz of `leaflet`; it's not `mapview`'s fault)

--

- but [`deck.gl`](https://deck.gl) will

--

- and it's available in R via the `mapdeck` package brought to us by 
`r icon::fa_github(animate = "spin")` `symbolixAU`

---

# Visualising osmdata

```{r promenade-data4, eval = FALSE}
promenade <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    add_osm_feature (key = "name", value = "promenade",
                     value_exact = FALSE) %>%
    osmdata_sf ()
```

---

# Visualising osmdata

A little bit of pre-processing #1: Because `mapdeck` is highly experimental and
currently needs to know where to centre the map.

```{r mapdeck-pre}
location <- ms48155$osm_lines %>%
    sf::st_coordinates () %>%
    apply (2, mean) %>%
    as.numeric () %>%
    magrittr::extract (1:2)
location
```

---

# Visualising osmdata

A little bit of pre-processing #2: Because `mapdeck` needs a personal mapbox
token which can be set with `set_token()`, and automatically used to generated
all maps

```{r mapdeck-token, eval = FALSE}
library (mapdeck)
set_token (<my_mapbox_token>)
```

---

# Visualising osmdata

```{r mapdeck3, eval = FALSE, echo = TRUE}
library (mapdeck)
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = location,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80) %>%
    add_path (data = promenade$osm_lines,
              layer_id = "promenade",
              stroke_width = 5,
              stroke_colour = "#33cc33",
              stroke_opacity = 100)
```
---
class: center, middle, inverse
background-image: url(img/ms-highway-promenade-mapview.png)
background-size: contain 
background-position: 50% 50%

---

# Visualising osmdata

```{r, eval = FALSE}
system.time (
             mapview (ms48155$osm_lines)
)
```
```{r, echo = FALSE, message = TRUE}
res <- c (3.799, 0.074, 3.906)
names (res) <- c ("user", "system", "elapsed")
res
```

```{r, eval = FALSE}
system.time (
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80)
)
```
```{r, echo = FALSE, message = TRUE}
res <- c (0.222, 0.000, 0.237)
names (res) <- c ("user", "system", "elapsed")
res
```

---

# Visualising osmdata

One further alternative for static plots is the [`osmplotr`
package](https://github.com/ropensci/osmplotr)

---
class: top, center, inverse

.fontBIG[
Question:<br><br> Has anybody <br>used `osmdata`?
]

---
class: top, center, inverse

.fontBIG[
Question:<br><br> What kind of default plot method might <br>be useful?
]

---

# osmdata

- The fastest and easiest method in any language for extracting OSM data
- Only software that allows OSM data to be extracted in 3 simple lines:

```{r, eval = FALSE}
ms <- opq ("münster de") %>%               # WHERE
    add_osm_feature (key = "highway") %>%  # WHAT
    osmdata_sf (quiet = FALSE)             # HOW
```

- Extremely flexible; can extract any form of OSM data.

```{r, eval = FALSE}
ms <- opq ("münster de") %>%
    add_osm_feature (key = "building") %>%
    osmdata_sf (quiet = FALSE)
```
- For more information, see [the website](https://github.com/ropensci/osmdata)

---
class: top, center, inverse

.fonthuge[
    **dodgr**<br>
    ]
.fontBIG[
    Distances on Directed GRaphs
]

---
class: top, center, inverse

.fonthuge[
    **dodgr**<br>
    ]

* CRAN <br><br>
* `r icon::fa_github(size = 4, animate = "spin", colour = "#ffcc00")` [`ATFutures/dodgr`](https://github.com/atfutures/dodgr)

---

# dodgr

Get all OSM highway data for Münster (not just 48155)

```{r osm-ms, eval = FALSE}
ms <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf ()
ms
```
```{r load-ms, echo = FALSE, eval = FALSE, message = TRUE}
ms <- readRDS ("data/osm-ms-highways.Rds")
ms
```
```{r load-ms-fakey, echo = FALSE, message = TRUE}
message ("Object of class 'osmdata' with:\n",
"                 $bbox : 51.8401448,7.4737853,52.0600251,7.7743634\n",
"        $overpass_call : The call submitted to the overpass API\n",
"                 $meta : metadata including timestamp and version numbers\n",
"           $osm_points : 'sf' Simple Features Collection with 157614 points\n",
"            $osm_lines : 'sf' Simple Features Collection with 33545 linestrings\n",
"         $osm_polygons : 'sf' Simple Features Collection with 510 polygons\n",
"       $osm_multilines : 'sf' Simple Features Collection with 0 multilinestrings\n",
"    $osm_multipolygons : 'sf' Simple Features Collection with 18 multipolygons")
```

---

# dodgr

Convert Münster OSM data to `dodgr` format

```{r weight-streetnet-ms1, eval = FALSE, message = TRUE}
ms <- osm_poly2line (ms) # osmdata function
```

---

# dodgr

Convert Münster OSM data to `dodgr` format

```{r weight-streetnet-ms3, eval = FALSE, message = TRUE}
ms <- osm_poly2line (ms) # osmdata function
library (dodgr)
net <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle") #<<
dim (net); head (net)
```
```{r weight-streetnet-ms-fakey, eval = TRUE, echo = FALSE, message = TRUE}
library (dodgr)
dim (net); head (net)
```

---

# dodgr distances

```{r ms-components, eval = TRUE, message = TRUE}
dim (net)
head (table (net$component), 10)
```

---

# dodgr distances

Calculate cycling distances between random points.

```{r dodgr-vertices2, eval = FALSE}
v <- dodgr_vertices (net)
dim (v); head (v)
```
```{r dodgr-vertices2-fakey, eval = TRUE, echo = FALSE, message = TRUE}
v <- dodgr_vertices (net)
dim (v); head (v, n = 5)
```
```{r vert-sample, eval = TRUE}
pts <- sample (v$id, size = 1000)
head (pts)
```

---

# dodgr distances

```{r dodgr-dists-help-fakey, echo = TRUE, eval = FALSE}
?dodgr_dists
```
```{r dodgr-dists-help, echo = FALSE, message = TRUE}
helpfile <- utils:::.getHelpFile (help ("dodgr_dists"))
hs <- capture.output (tools:::Rd2txt(helpfile))
#hs <- capture.output (tools:::prepare_Rd(helpfile))
cat(hs, sep="\n")
```

---

# dodgr distances

Calculate cycling distances between random points.

```{r dodgr-vertices3, eval = FALSE}
v <- dodgr_vertices (net)
dim (v); head (v)
```
```{r dodgr-vertices3-fakey, eval = TRUE, echo = FALSE, message = TRUE}
dim (v); head (v, n = 5)
```
```{r vert-sample3, eval = TRUE, message = TRUE}
pts <- sample (v$id, size = 1000)
head (pts)
```

---

# dodgr distances

```{r dodgr-dists1-copy, eval = FALSE}
system.time (
             d <- dodgr_dists (net, from = pts, to = pts) #<<
             )
```
```{r dodgr-dists1-copy-fakey, echo = FALSE, message = TRUE}
message ("   user  system elapsed\n",
" 43.847   0.013   6.042")
```

--

```{r dodgr-dists-result1, eval = FALSE}
class (d); dim (d); range (d); range (d, na.rm = TRUE)
```
```{r dodgr-dists-result1-fakey, eval = TRUE, echo = FALSE, message = TRUE}
message ("[1] matrix\n",    # explicit print coz otherwise output inserts
         "[1] 1000 1000\n", # blank lines that no longer fit on screen
         "[1] NA NA\n",
         "[1] 0.00000 35.50599")
```
```{r dodgr-dists-result2, eval = FALSE}
round (1000 ^ 2 * length (which (is.na (d))) / length (d))
```
```{r dodgr-dists-result2-fakey, eval = TRUE, echo = FALSE, message = TRUE}
message ("[1] 32709")
```

---

# dodgr distances

```{r dodgr-dists1, eval = FALSE}
system.time (
             d <- dodgr_dists (net, from = pts, to = pts) #<<
             )
```
```{r dodgr-dists1-fakey, echo = FALSE, message = TRUE}
message ("   user  system elapsed\n",
" 43.847   0.013   6.042")
```

- Those 1,000,000 distances could have been calculated with the google API
- And that would cost `1e6 * 0.004 = $US4,000`
- $US 4,000 for 6 seconds of calculation!
- To date, `dodgr` provides one of the only .bold[free] ways to perform such
  calculations, and is also entirely .bold[open source].

---
class: left, top, inverse
background-image: url(img/ms-promenade.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

---

class: inverse, center, top

.fonthuge[
    ~~distances~~<br><br>
    flows
    ]

---

# flows

If one person travels between each of our 1,000,000 pairs of points, how many
travel on each bit of the network?

```{r, eval = FALSE}
d <- dodgr_dists (net, from = pts, to = pts)
```
```{r dodgr-flows, eval = FALSE}
flow_mat <- matrix (1, nrow = 1000, ncol = 1000)
system.time (
    f <- dodgr_flows_aggregate (net, from = pts, to = pts,  #<<
                                flows = flow_mat)           #<<
)
```
```{r, eval = FALSE, echo = FALSE}
saveRDS (f, file = "data/dodgr-flows-ms.Rds")
```

```{r dodgr-flows-fakey, eval = TRUE, echo = FALSE, message = TRUE}
message ("   user  system elapsed\n",
"145.186   1.107  20.016")
```

---

# flows

```{r dodgr-flows-result-d, eval = TRUE, message = TRUE}
head (net)
```

---

# flows

```{r dodgr-flows-result, eval = TRUE, message = TRUE}
head (f)
```

---

# flows

```{r merge-directed-flows, eval = TRUE, message = TRUE}
fm <- merge_directed_flows (f)
```

- Adds flows .bold[A] `r icon::fa_long_arrow_alt_right()` .bold[B] to .bold[B]
  `r icon::fa_long_arrow_alt_right()` .bold[A]
- Removes one of the two edges
- Also removes all edges with zero flow

--

```{r merge-directed-flows-result, message = TRUE}
dim (f); dim (fm)
nrow (fm) / nrow (f)
```

---

# flows

```{r dodgr-flows-result2, eval = TRUE, message = TRUE}
head (fm)
```

---

# flows

```{r flowplot, out.height = "400px", out.width = "500px", echo = TRUE}
library (ggplot2)
ggplot (fm, aes (flow)) + geom_freqpoly ()
```

---
class: center, inverse

.fonthuge[
mapdeck <br> time!<br>
`r icon::fa_uber(animate = "spin")`
]
---

# flows

```{r ms-mapdeck, eval = FALSE}
library (mapdeck)
f$flow <- 20 * f$flow / max (f$flow)
pal <- colorRampPalette (c ("lawngreen", "red"))
loc <- apply (sf::st_coordinates (ms$osm_lines), 2, mean) [1:2]
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         zoom = 12,
         location = loc) %>%
    add_line (data = fm,
              layer_id = "ms-highways",
              origin = c("from_lon", "from_lat"),
              destination = c("to_lon", "to_lat"),
              stroke_colour = "flow",
              stroke_width = "flow",
              palette = pal)
```
```{r ms-mapdeck-fakey, echo = FALSE}
loc <-  c (7.619786, 51.952610)
```

---
background-image: url(img/ms-random-flows.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade100.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
class: center, inverse

.fontBIG[
The Promenade<br>IS<br>cycling in Münster,<br>but there are no flows along most of
it!
]

---

# How to get people riding along the Promenade?

- We first need to find the edges of our `dodgr net` which correspond to the
Promenade.
- Get OSM way IDs (making sure to exclude the "Kanalpromenade")

```{r promenade-ids, eval = FALSE}
index <- grep ("promenade", ms$osm_lines$name,
               ignore.case = TRUE)
index_kp <- grep ("kanal", ms$osm_lines$name,
                  ignore.case = TRUE)
index <- index [!index %in% index_kp]
ids_prom <- ms$osm_lines$osm_id [index] %>%
    as.character ()
```

---

# How to get people riding along the Promenade?

- Presume the Promenade "feels" 10% shorter than what it actually is

```{r promenade-reweight, eval = FALSE}
shorter <- 0.1
g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
index <- which (g$way_id %in% ids_prom)
g$d_weighted [index] <- g$d [index] * (1 - shorter) #<<
f <- dodgr_flows_aggregate (g, from = pts, to = pts,
                            flows = flowmat)
```

---

# How to get people riding along the Promenade?

```{r}
head (net)
```

---

# How to get people riding along the Promenade?

- Presume the Promenade "feels" 10% shorter than what it actually is

```{r promenade-reweight2, eval = FALSE}
shorter <- 0.1
g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
index <- which (g$way_id %in% ids_prom)
g$d_weighted [index] <- g$d [index] * (1 - shorter)
f <- dodgr_flows_aggregate (g, from = pts, to = pts,
                            flows = flowmat)
```

---
background-image: url(img/ms-promenade90.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---

# How to get people riding along the Promenade?

Presume the Promenade "feels" (0, 10, 20, & 30)% shorter than what it actually is

```{r promenade-reweight3, eval = FALSE}
g <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle")
index <- which (g$way_id %in% ids)
g$d_weighted [index] <- g$d [index] * (1 - shorter)
f <- dodgr_flows_aggregate (g, from = pts, to = pts,
                            flows = flowmat)
```

---
background-image: url(img/ms-promenade100.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade90.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade80.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade70.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
class: center, top, inverse
.fontBIG[
The Promenade feels like it is 30% shorter than what it actually is
]

---
class: center, top, inverse
.fontBIG[
Münsteraner_Innen are willing to ride 30% further to ride along the Promenade
]

---
class: center, top, inverse
.fontBIG[
Important Lesson: Cultural specificities are important!
]

--

.fontBIG[
And can be incorporated!
]

---
class: center, top, inverse
.fontBIG[
Popquiz: <br><br>Where in Münster are cyclists most exposed to air pollution?
]

---
background-image: url(img/ms-bike-flows.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---

# Where is Münster are cyclists most exposed to air pollution?

- Air pollution comes from automotive vehicles, so need to know where they
  travel

- Generate another flow layer

--

- Implement model to simulate dispersal to adjacent ways including bike paths

---
background-image: url(img/ms-car-flows.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-car-flows-dispersed.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade70.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---

# Where is Münster are cyclists most exposed to air pollution?

- Air pollution comes from automotive vehicles, so need to know where they
  travel

- Generate another flow layer

- Implement model to simulate dispersal to adjacent ways including bike paths

- Quantify exposure as cyclist flows times pollutant concentrations

---
background-image: url(img/ms-exposure.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-car-flows-dispersed.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-promenade70.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

---
background-image: url(img/ms-exposure.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000
class: center, middle, inverse

--

.fonthuge[THANK<br>YOU!]
.box-bottom[May the flow be with you]

---
class: center, middle, inverse
background-image: url(img/ms-exposure.png)
background-size: contain 
background-position: 50% 50%
background-color: #000000

.left-column[
`r icon::fa_github(size = 2, animate = "spin")` atfutures

`r icon::fa_github(size = 2, animate = "spin")` mpadge

`r icon::fa_github(size = 2, animate = "spin")` ropensci
]

.right-column[
`r icon::fa_twitter(size = 2, animate = "spin")` bikesRdata

`r icon::fa_envelope(size = 1)` .small[mark.padgham@email.com]<br><br>

`r icon::ii_ios_world_outline(size = 5)` atfutures.github.io
]

.box-bottom[
slides at <br>
[https://github.com/mpadge/ms-user-meetup-nov2018](https://github.com/mpadge/ms-user-meetup-nov2018)
]